Towards a standardized evaluation of imputation methodology

### 2022-07-27 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Fix referencing issues, add check boxes to check list

Standardizing the evaluation of imputation methodology requires simulators to consider common aspects in their simulation workflows. Table \ref{table:steps} outlines some suggested steps to adopt. Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach. Our suggestions are the accumulated result of scientific literature, research experience, and discussion. For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}. We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology. [TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]

@@

\begin{table}[ht]
\caption{Checklist for reporting on imputation methodology evaluations.}
\label{table:check}
\begin{tabular}{ll}
% \hline
%   & Suggested aspects to report                                                       \\
\hline
1 & Simulation scope                                                                   \\ \cline{2-2}
  & Aim                                                                                \\
  & Design (incl. pseudo-code or flow diagram)                                         \\
  & Number of simulation repetitions                                                   \\ \hline
2 & Comparative truth                                                                  \\ \cline{2-2}
  & Data source (model-based or design-based)                                           \\
  & Sampling variance                                                                  \\
%  & Data characteristics (number of observations, number of variables, variable types, \\
%  & coherence between variables incl. data structures e.g. clustering)                 \\
  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
  & Estimand                                                                           \\ \hline 
3 & Induced missingness                                                                \\ \cline{2-2}
  & Missingness mechanism (incl. type or functional form of the missing data   model)  \\
  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
4 & Applied methods                                                                    \\ \cline{2-2}
  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
  & Analytic methods (incl. calculation of standard errors e.g. pooling   rules)       \\
  & Reference method (e.g. complete case analysis)                                    \\ \hline 
5 & Imputation evaluation                                                              \\ \cline{2-2}
  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
  & Imputation model fit (e.g. posterior predictive checks)                            \\
  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
6 & Performance evaluation                                                             \\ \cline{2-2}
  & Statistical properties (e.g. confidence validity)                                  \\
  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
\end{tabular}
\end{table}

>>>>

Standardizing the evaluation of imputation methodology requires simulators to consider common aspects in their simulation workflows. Table \ref{table:check} outlines some suggested steps to adopt. Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach. Our suggestions are the accumulated result of scientific literature, research experience, and discussion. For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}. We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology. [TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]

@@

\begin{table}[ht]
\caption{Checklist for reporting on imputation methodology evaluations.}
\label{table:check}
\begin{tabular}{ll}
% \hline
%   & Suggested aspects to report                                                       \\
\hline
1 & Simulation scope                                                                   \\ \cline{2-2}
$\square$  & Aim                                                                                \\
$\square$  & Design (incl. pseudo-code or flow diagram)                                         \\
$\square$  & Number of simulation repetitions                                                   \\ \hline
2 & Comparative truth                                                                  \\ \cline{2-2}
$\square$  & Data source (model-based or design-based)                                           \\
$\square$  & Sampling variance                                                                  \\
%  & Data characteristics (number of observations, number of variables, variable types, \\
%  & coherence between variables incl. data structures e.g. clustering)                 \\
$\square$  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
$\square$  & Estimand                                                                           \\ \hline 
3 & Induced missingness                                                                \\ \cline{2-2}
$\square$  & Missingness mechanism (incl. type or functional form of the missing data   model)  \\
$\square$  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
4 & Applied methods                                                                    \\ \cline{2-2}
$\square$  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
$\square$  & Analytic methods (incl. calculation of standard errors e.g. pooling   rules)       \\
$\square$  & Reference method (e.g. complete case analysis)                                    \\ \hline 
5 & Imputation evaluation                                                              \\ \cline{2-2}
$\square$  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
$\square$  & Imputation model fit (e.g. posterior predictive checks)                            \\
$\square$  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
6 & Performance evaluation                                                             \\ \cline{2-2}
$\square$  & Statistical properties (e.g. confidence validity)                                  \\
$\square$  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
\end{tabular}
\end{table}
